---
title: "Seasonal variation in the hydra microbiome"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r message=FALSE, warning=FALSE}
library(phyloseq); library(ggplot2); library(metagMisc); library(fantaxtic);
library(ANCOMBC); library(pairwiseAdonis); library(dplyr); library(microViz);

rm(list=ls())
load("seasonal_phyloseq.Rdata")
tre=ape::read.tree("/storage/microbiome/seasonal_2024/full_tree_midpointRooted.tre")
phy_tree(ps_seasonal) <- tre

ps_seasonal <- subset_taxa(ps_seasonal, (Order!="Chloroplast")|(!is.na(Class)))
ps_seasonal <- subset_taxa(ps_seasonal, (Genus!="unidentified_Chloroplast")|is.na(Genus))
ps_seasonal <- subset_taxa(ps_seasonal, (Genus!="unidentified_Mitochondria")|is.na(Genus))
ps_seasonal <- subset_taxa(ps_seasonal, (Family!="Mitochondria")|is.na(Family))

ps_seasonal <- name_na_taxa(ps_seasonal, include_rank = F, na_label="Unknown <tax>")
ps_seasonal <- tax_glom(ps_seasonal, "Genus")
ps_seasonal <- label_duplicate_taxa(ps_seasonal, tax_level="Genus")
taxa_names(ps_seasonal) <- phyloseq::tax_table(ps_seasonal)[,6]

ps_seasonal <- subset_samples(ps_seasonal, sample_sums(ps_seasonal) > 5000)
ps_seasonal <- phyloseq_filter_prevalence(ps_seasonal, prev.trh = 0.05)
ps_seasonal <- phyloseq_filter_taxa_tot_fraction(ps_seasonal, frac=0.00005)

sample_data(ps_seasonal)$Sampling <- factor(sample_data(ps_seasonal)$Sampling,
                                            levels=c("2021 June",
                                                     "2021 September",
                                                     "2021 December",
                                                     "2022 March",
                                                     "2022 June",
                                                     "2022 November",
                                                     "2023 February",
                                                     "2023 April",
                                                     "2023 June"))
sample_data(ps_seasonal)$Year <- ifelse(sample_data(ps_seasonal)$Sampling %in% 
                                          c("2021 June", "2021 September", "2021 December",
                                            "2022 March", "2022 June"), "2021/22", "2022/23")

sample_data(ps_seasonal)$Season <- factor(sample_data(ps_seasonal)$Season, 
                                          levels=c("Spring","Summer","Autumn","Winter"))

sample_data(ps_seasonal)$PopID <- factor(sample_data(ps_seasonal)$PopID,
                                         levels=c("M52","M25","M26"))
levels(sample_data(ps_seasonal)$PopID) <- c("Shallow lake","Deep lake","River")

sample_data(ps_seasonal)$SiteID <- 
  gsub("W","",gsub("[0-9]*$","",row.names(sample_data(ps_seasonal))))

summary(sample_sums(ps_seasonal))
table(sample_data(ps_seasonal)$Sampling, sample_data(ps_seasonal)$Type,
                                              sample_data(ps_seasonal)$PopID)

ps_seasonal <- rarefy_even_depth(ps_seasonal)
```

```{r}
library(microbiome); library(RColorBrewer)

prevalences <- seq(.05, 1, .2)
detections <- round(10^seq(log10(0.002), log10(.2), length = 9), 3)

polyp_M25_rel <- microbiome::transform(
  subset_samples(ps_seasonal, Type=="polyp"&PopID=="Deep lake"), "compositional")
polyp_M26_rel <- microbiome::transform(
  subset_samples(ps_seasonal, Type=="polyp"&PopID=="River"), "compositional")
polyp_M52_rel <- microbiome::transform(
  subset_samples(ps_seasonal, Type=="polyp"&PopID=="Shallow lake"), "compositional")


pcl.M25 <- plot_core(polyp_M25_rel, 
          prevalences = prevalences, 
          detections = c(0, 0.1, 0.5, 2, 5, 20)/100, 
          plot.type = "lineplot") + 
  xlab("Relative Abundance (%)")
pcl.M26 <- plot_core(polyp_M26_rel, 
          prevalences = prevalences, 
          detections = c(0, 0.1, 0.5, 2, 5, 20)/100, 
          plot.type = "lineplot") + 
  xlab("Relative Abundance (%)")
pcl.M52 <- plot_core(polyp_M52_rel, 
          prevalences = prevalences, 
          detections = c(0, 0.1, 0.5, 2, 5, 20)/100, 
          plot.type = "lineplot") + 
  xlab("Relative Abundance (%)")
ggpubr::ggarrange(pcl.M52, pcl.M25, pcl.M26, common.legend=T)


pchm.M52 <- plot_core(polyp_M52_rel,horizontal=T,
               plot.type = "heatmap", 
               colours = rev(brewer.pal(5, "Spectral")),
               prevalences = prevalences, 
               detections = detections, 
               min.prevalence = prevalence(polyp_M25_rel, sort = TRUE)[100]) +
  labs(x = "Detection Threshold\n(Relative Abundance (%))") +
  scale_fill_gradientn(
      colours = rev(brewer.pal(9, "Spectral")),  # Specify the color palette
      name = "Prevalence:"                      # Optional: add a legend title
  ) +
  theme(axis.text.y= element_text(size=8, face="italic"),
        axis.text.x.bottom=element_text(size=9,angle=45,hjust = 1),
        axis.title = element_text(size=10),
        legend.text = element_text(size=8),
        legend.title = element_text(size=10))+labs(title="Shallow lake")

pchm.M25 <- plot_core(polyp_M25_rel,horizontal=T,
               plot.type = "heatmap", 
               colours = rev(brewer.pal(5, "Spectral")),
               prevalences = prevalences, 
               detections = detections, 
               min.prevalence = prevalence(polyp_M25_rel, sort = TRUE)[100]) + labs(x="")+
  #labs(x = "Detection Threshold\n(Relative Abundance (%))") +
  theme(axis.text.y= element_text(size=8, face="italic"),
        axis.text.x.bottom=element_text(size=9,angle=45,hjust = 1),
        axis.title = element_text(size=10),
        legend.text = element_text(size=8),
        legend.title = element_text(size=10))+labs(title="Deep lake")

pchm.M26 <- plot_core(polyp_M26_rel,horizontal=T,
               plot.type = "heatmap", 
               colours = rev(brewer.pal(5, "Spectral")),
               prevalences = prevalences, 
               detections = detections, 
               min.prevalence = prevalence(polyp_M25_rel, sort = TRUE)[100]) + labs(x="")+
  theme(axis.text.y= element_text(size=8, face="italic"),
        axis.text.x.bottom=element_text(size=9,angle=45,hjust = 1),
        axis.title = element_text(size=10),
        legend.text = element_text(size=8),
        legend.title = element_text(size=10))+labs(title="River")

widths <- c(length(levels(pchm.M52$data$Taxa)),
                               length(levels(pchm.M25$data$Taxa)),
                               length(levels(pchm.M26$data$Taxa)))

ragg::agg_tiff(filename="~/hidra/2023/MicrobiomeSeasonal/MS/figs/polyp_core.tiff",
               width=10, height=6,units="in",res=300, scaling=0.75)
ggpubr::ggarrange(pchm.M52, pchm.M25, pchm.M26, common.legend=T,ncol=3,align="h",
                  widths=widths/sum(widths))
dev.off()
```

```{r}
library(microbiome); library(RColorBrewer)

prevalences <- seq(0.1, 1, .2)
detections <- round(10^seq(log10(0.006), log10(.2), length = 9), 3)

water_M25_rel <- microbiome::transform(
  subset_samples(ps_seasonal, Type=="water"&PopID=="Deep lake"), "compositional")
water_M26_rel <- microbiome::transform(
  subset_samples(ps_seasonal, Type=="water"&PopID=="River"), "compositional")
water_M52_rel <- microbiome::transform(
  subset_samples(ps_seasonal, Type=="water"&PopID=="Shallow lake"), "compositional")


pcl.M25 <- plot_core(water_M25_rel, 
          prevalences = prevalences, 
          detections = c(0, 0.1, 0.5, 2, 5, 20)/100, 
          plot.type = "lineplot") + 
  xlab("Relative Abundance (%)")
pcl.M26 <- plot_core(water_M26_rel, 
          prevalences = prevalences, 
          detections = c(0, 0.1, 0.5, 2, 5, 20)/100, 
          plot.type = "lineplot") + 
  xlab("Relative Abundance (%)")
pcl.M52 <- plot_core(water_M52_rel, 
          prevalences = prevalences, 
          detections = c(0, 0.1, 0.5, 2, 5, 20)/100, 
          plot.type = "lineplot") + 
  xlab("Relative Abundance (%)")
ggpubr::ggarrange(pcl.M52, pcl.M25, pcl.M26, common.legend=T)


pchm.M52 <- plot_core(water_M52_rel,horizontal=T,
               plot.type = "heatmap", 
               colours = rev(brewer.pal(5, "Spectral")),
               prevalences = prevalences, 
               detections = detections, 
               min.prevalence = prevalence(polyp_M25_rel, sort = TRUE)[100]) +
  labs(x = "Detection Threshold\n(Relative Abundance (%))") +
  scale_fill_gradientn(
      colours = rev(brewer.pal(9, "Spectral")),  # Specify the color palette
      name = "Prevalence:"                      # Optional: add a legend title
  ) +
  theme(axis.text.y= element_text(size=8, face="italic"),
        axis.text.x.bottom=element_text(size=9,angle=45,hjust = 1),
        axis.title = element_text(size=10),
        legend.text = element_text(size=8),
        legend.title = element_text(size=10))+labs(title="Shallow lake")

pchm.M25 <- plot_core(water_M25_rel,horizontal=T,
               plot.type = "heatmap", 
               colours = rev(brewer.pal(5, "Spectral")),
               prevalences = prevalences, 
               detections = detections, 
               min.prevalence = prevalence(polyp_M25_rel, sort = TRUE)[100]) + labs(x="")+
  #labs(x = "Detection Threshold\n(Relative Abundance (%))") +
  theme(axis.text.y= element_text(size=8, face="italic"),
        axis.text.x.bottom=element_text(size=9,angle=45,hjust = 1),
        axis.title = element_text(size=10),
        legend.text = element_text(size=8),
        legend.title = element_text(size=10))+labs(title="Deep lake")

pchm.M26 <- plot_core(water_M26_rel,horizontal=T,
               plot.type = "heatmap", 
               colours = rev(brewer.pal(5, "Spectral")),
               prevalences = prevalences, 
               detections = detections, 
               min.prevalence = prevalence(polyp_M25_rel, sort = TRUE)[100]) + labs(x="")+
  theme(axis.text.y= element_text(size=8, face="italic"),
        axis.text.x.bottom=element_text(size=9,angle=45,hjust = 1),
        axis.title = element_text(size=10),
        legend.text = element_text(size=8),
        legend.title = element_text(size=10))+labs(title="River")

widths <- c(length(levels(pchm.M52$data$Taxa)),
                               length(levels(pchm.M25$data$Taxa)),
                               length(levels(pchm.M26$data$Taxa)))

ragg::agg_tiff(filename="~/hidra/2023/MicrobiomeSeasonal/MS/figs/water_core.tiff",
               width=10, height=6,units="in",res=300, scaling=0.75)
ggpubr::ggarrange(pchm.M52, pchm.M25, pchm.M26, common.legend=T,ncol=3,align="h",
                  widths=widths/sum(widths))
dev.off()
```


# Alpha diversity calculations & plots
```{r alpha1, fig.height=10, fig.width=10, message=F, warning=F}
sample_data(ps_seasonal)$chao1 <- microbiome::alpha(ps_seasonal, "chao1")[,1]
sample_data(ps_seasonal)$faith_pd <- btools::estimate_pd(ps_seasonal)$PD
sample_data(ps_seasonal)$shannon <- microbiome::alpha(ps_seasonal, "shannon")[,1]
sample_data(ps_seasonal)$observed <- microbiome::alpha(ps_seasonal, "observed")[,1]

alpha_plot <- ggplot(sample_data(ps_seasonal), aes(y=chao1, x=Sampling, fill=Season))+
  geom_boxplot()+facet_wrap(~PopID*Type,ncol=2)+theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  labs(title="Alpha diversity")+
  ylab("Chao1 diversity index")+
  scale_fill_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))
alpha_plot2 <- ggplot(sample_data(ps_seasonal), aes(y=faith_pd, x=Sampling, fill=Season))+
  geom_boxplot()+facet_wrap(~PopID*Type,ncol=2)+theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  labs(title="Alpha diversity")+
  ylab("Faith's phylogenetic diversity index")+
  scale_fill_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))
alpha_plot3 <- ggplot(sample_data(ps_seasonal), aes(y=shannon, x=Sampling, fill=Season))+
  geom_boxplot()+facet_wrap(~PopID*Type,ncol=2)+theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  labs(title="Alpha diversity")+
  ylab("Shannon diversity index")+
  scale_fill_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))
alpha_plot4 <- ggplot(sample_data(ps_seasonal), aes(y=observed, x=Sampling, fill=Season))+
  geom_boxplot()+facet_wrap(~PopID*Type,ncol=2)+theme_bw()+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  labs(title="Alpha diversity")+
  ylab("No. of observed features")+
  scale_fill_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))

ragg::agg_tiff(filename="/storage/microbiome/seasonal_2024/figs/alpha_div.tiff",
               width=10, height=10,units="in",res=300, scaling=0.75)
ggpubr::ggarrange(alpha_plot+labs(title="Chao1 index"), 
    #              alpha_plot2+labs(title="Faith's PD"), 
                  alpha_plot3+labs(title="Shannon index"), 
                  alpha_plot4+labs(title="Observed features"), ncol=2, nrow=2,
                  labels="AUTO",common.legend=T)
dev.off()

polyp_div <- sample_data(ps_seasonal)%>%data.frame()%>%filter(Type=="polyp")%>%
  select("SeqID","PopID","Sampling","Season","chao1","faith_pd","shannon","observed")
water_div <- sample_data(ps_seasonal)%>%data.frame()%>%filter(Type=="water")%>%
  select("SeqID","PopID","Sampling","Temp","chao1","faith_pd","shannon","observed")

polyp_div$SiteID <- gsub("[0-9]*$","",polyp_div$SeqID) 
water_div$SiteID <- gsub("W","",water_div$SeqID)

names(polyp_div) <- paste("polyp", names(polyp_div), sep="_")
names(water_div) <- paste("water", names(water_div), sep="_")

div <- left_join(polyp_div, water_div, by=join_by("polyp_SiteID"=="water_SiteID"))
```

# Beta diversity calculations & plots
```{r, fig.width=10, message=F, warning=F}
siteall_comp <- microbiome::transform(ps_seasonal,"identity")
nmds.obj_all.bray <- ordinate(siteall_comp,method="PCoA", distance="bray")
beta.braycurtis_all <- plot_ordination(siteall_comp,nmds.obj_all.bray,
                                    type="samples",color="Season",shape="Type")+
  theme_bw()+geom_point(size=3)+facet_wrap(~PopID)+
  labs(title="Beta diversity - Bray Curtis")+
  scale_color_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))

beta.braycurtis_all

nmds.obj_all.jaccard <- ordinate(siteall_comp,method="PCoA", distance="jaccard", binary=T)
beta.jaccard_all <- plot_ordination(siteall_comp,nmds.obj_all.jaccard,
                                    type="samples",color="Season",shape="Type")+
  theme_bw()+geom_point(size=3)+facet_wrap(~PopID)+
  labs(title="Beta diversity - Jaccard presence-absence")+
  scale_color_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))
beta.jaccard_all

```

```{r}
library(ggpubr)
ragg::agg_tiff(filename="/storage/microbiome/seasonal_2024/figs/ab_div.tiff",
               width=8, height=6,units="in",res=300, scaling=0.8)
ggarrange(alpha_plot+theme(legend.position = "bottom",
                legend.background = element_rect(color = "black", linewidth = 0.25)), 
          beta.braycurtis_all+
  guides(
    color = guide_legend(order = 2), # Set order of legends
    shape = guide_legend(order = 1)
  ) +
  theme(
    legend.position="bottom",
    legend.box = "vertical",         # Stack legends vertically
    legend.spacing.y = unit(0.0, 'cm'), # Adjust vertical spacing between legends
    legend.background = element_rect(color = "black", linewidth = 0.25)   #Draw box around each legend
    ),
  labels="AUTO")
dev.off()

```

```{r}
sample_data(siteall_comp)$TypeSeason <- interaction(sample_data(siteall_comp)$Type,sample_data(siteall_comp)$Season)
sample_data(siteall_comp)$Site <- gsub("víz/","",sample_data(siteall_comp)$SamplingName)
sample_data(siteall_comp)$Site <- gsub("/[0-9]*","",sample_data(siteall_comp)$Site)
sample_data(siteall_comp)$Site <- paste(sample_data(siteall_comp)$Sampling, sample_data(siteall_comp)$Site)

samplesM25 <- subset_samples(siteall_comp, PopID=="Deep lake")
samplesM26 <- subset_samples(siteall_comp, PopID=="River")
samplesM52 <- subset_samples(siteall_comp, PopID=="Shallow lake")

polypsM25 <- subset_samples(siteall_comp, PopID=="Deep lake"&Type=="polyp")
polypsM26 <- subset_samples(siteall_comp, PopID=="River"&Type=="polyp")
polypsM52 <- subset_samples(siteall_comp, PopID=="Shallow lake"&Type=="polyp")

waterM25 <- subset_samples(siteall_comp, PopID=="Deep lake"&Type=="water")
waterM26 <- subset_samples(siteall_comp, PopID=="River"&Type=="water")
waterM52 <- subset_samples(siteall_comp, PopID=="Shallow lake"&Type=="water")


dist.bray.M25 <- phyloseq::distance(samplesM25, method="bray")
dist.bray.M26 <- phyloseq::distance(samplesM26, method="bray")
dist.bray.M52 <- phyloseq::distance(samplesM52, method="bray")

dist.bray.M25.polyps <- phyloseq::distance(polypsM25, method="bray")
dist.bray.M26.polyps <- phyloseq::distance(polypsM26, method="bray")
dist.bray.M52.polyps <- phyloseq::distance(polypsM52, method="bray")

dist.bray.M25.water <- phyloseq::distance(waterM25, method="bray")
dist.bray.M26.water <- phyloseq::distance(waterM26, method="bray")
dist.bray.M52.water <- phyloseq::distance(waterM52, method="bray")

dist.jaccard.M25 <- phyloseq::distance(samplesM25, method="jaccard", binary=T)
dist.jaccard.M26 <- phyloseq::distance(samplesM26, method="jaccard", binary=T)
dist.jaccard.M52 <- phyloseq::distance(samplesM52, method="jaccard", binary=T)

dist.jaccard.M25.polyps <- phyloseq::distance(polypsM25, method="jaccard", binary=T)
dist.jaccard.M26.polyps <- phyloseq::distance(polypsM26, method="jaccard", binary=T)
dist.jaccard.M52.polyps <- phyloseq::distance(polypsM52, method="jaccard", binary=T)

dist.jaccard.M25.water <- phyloseq::distance(waterM25, method="jaccard", binary=T)
dist.jaccard.M26.water <- phyloseq::distance(waterM26, method="jaccard", binary=T)
dist.jaccard.M52.water <- phyloseq::distance(waterM52, method="jaccard", binary=T)

```


# Compare diversity of water and polyp samples

```{r}
library(glmmTMB); library(MuMIn)
div.res <- data.frame(Pop=c("Shallow lake","Deep lake","River"),
                      chao1_p=NA, chao1_R2=NA,
                      shannon_p=NA, shannon_R2=NA,
                      obsfeat_p=NA, obsfeat_R2=NA,
                      BC_p=NA, BC_R2=NA,
                      Jac_p=NA, Jac_R2=NA)

alpha_type_m52 <- glmmTMB(chao1~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="Shallow lake"))
div.res$chao1_p[1] <- round(anova(alpha_type_m52, update(alpha_type_m52, .~.-Type))[[8]][2],3)
div.res$chao1_R2[1] <- r.squaredGLMM(alpha_type_m52)[1]

alpha_type_m25 <- glmmTMB(chao1~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="Deep lake"))
div.res$chao1_p[2] <- round(anova(alpha_type_m25, update(alpha_type_m25, .~.-Type))[[8]][2],3)
div.res$chao1_R2[2] <- r.squaredGLMM(alpha_type_m25)[1]

alpha_type_m26 <- glmmTMB(chao1~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="River"))
div.res$chao1_p[3] <- round(anova(alpha_type_m26, update(alpha_type_m26, .~.-Type))[[8]][2],3)
div.res$chao1_R2[3] <- r.squaredGLMM(alpha_type_m26)[1]

alpha_type_m52 <- glmmTMB(shannon~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="Shallow lake"))
div.res$shannon_p[1] <- round(anova(alpha_type_m52, update(alpha_type_m52, .~.-Type))[[8]][2],3)
div.res$shannon_R2[1] <- r.squaredGLMM(alpha_type_m52)[1]

alpha_type_m25 <- glmmTMB(shannon~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="Deep lake"))
div.res$shannon_p[2] <- round(anova(alpha_type_m25, update(alpha_type_m25, .~.-Type))[[8]][2],3)
div.res$shannon_R2[2] <- r.squaredGLMM(alpha_type_m25)[1]

alpha_type_m26 <- glmmTMB(shannon~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="River"))
div.res$shannon_p[3] <- round(anova(alpha_type_m26, update(alpha_type_m26, .~.-Type))[[8]][2],3)
div.res$shannon_R2[3] <- r.squaredGLMM(alpha_type_m26)[1]

alpha_type_m52 <- glmmTMB(observed~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="Shallow lake"))
div.res$obsfeat_p[1] <- round(anova(alpha_type_m52, update(alpha_type_m52, .~.-Type))[[8]][2],3)
div.res$obsfeat_R2[1] <- r.squaredGLMM(alpha_type_m52)[1]

alpha_type_m25 <- glmmTMB(observed~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="Deep lake"))
div.res$obsfeat_p[2] <- round(anova(alpha_type_m25, update(alpha_type_m25, .~.-Type))[[8]][2],3)
div.res$obsfeat_R2[2] <- r.squaredGLMM(alpha_type_m25)[1]

alpha_type_m26 <- glmmTMB(observed~Type,
                          data=sample_data(ps_seasonal)%>%data.frame()%>%
                            filter(PopID=="River"))
div.res$obsfeat_p[3] <- round(anova(alpha_type_m26, update(alpha_type_m26, .~.-Type))[[8]][2],3)
div.res$obsfeat_R2[3] <- r.squaredGLMM(alpha_type_m26)[1]

ad.M25 <- vegan::adonis2(dist.bray.M25~Type, data=data.frame(sample_data(samplesM25)),
                         by="onedf")
ad.M26 <- vegan::adonis2(dist.bray.M26~Type, data=data.frame(sample_data(samplesM26)),
                         by="onedf")
ad.M52 <- vegan::adonis2(dist.bray.M52~Type, data=data.frame(sample_data(samplesM52)),
                         by="onedf")
div.res$BC_p[1] <- ad.M52[[5]][1]
div.res$BC_R2[1] <- ad.M52[[3]][1]
div.res$BC_p[2] <- ad.M25[[5]][1]
div.res$BC_R2[2] <- ad.M25[[3]][1]
div.res$BC_p[3] <- ad.M26[[5]][1]
div.res$BC_R2[3] <- ad.M26[[3]][1]

ad.M25 <- vegan::adonis2(dist.jaccard.M25~Type, data=data.frame(sample_data(samplesM25)),
                         by="onedf")
ad.M26 <- vegan::adonis2(dist.jaccard.M26~Type, data=data.frame(sample_data(samplesM26)),
                         by="onedf")
ad.M52 <- vegan::adonis2(dist.jaccard.M52~Type, data=data.frame(sample_data(samplesM52)),
                         by="onedf")
div.res$Jac_p[1] <- ad.M52[[5]][1]
div.res$Jac_R2[1] <- ad.M52[[3]][1]
div.res$Jac_p[2] <- ad.M25[[5]][1]
div.res$Jac_R2[2] <- ad.M25[[3]][1]
div.res$Jac_p[3] <- ad.M26[[5]][1]
div.res$Jac_R2[3] <- ad.M26[[3]][1]
```

# Predictors of variation in alpha and beta diversity (season, temperature & bacterioplankton diversity)

```{r, fig.height=10, fig.width=10, message=F, warning=F}
library(glmmTMB); library(DHARMa); library(MuMIn)

###################################################################################
##################################### Shallow lake ################################
x00 <- glmmTMB(polyp_chao1~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"))
x1 <- glmmTMB(polyp_chao1~scale(water_Temp)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"))
x2 <- glmmTMB(polyp_chao1~polyp_Season+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"))
x01 <- glmmTMB(water_chao1~1, 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)))
x3 <- glmmTMB(water_chao1~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)))
x4 <- glmmTMB(water_chao1~polyp_Season, 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)))
x5 <- glmmTMB(polyp_chao1~scale(water_chao1)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"))

extract_info <- function(model,nullmodel) {
  pvals <- anova(model, nullmodel)[[8]][2]
  r2m <- r.squaredGLMM(model)[1]
  return(list(R2m = r2m, pvals = pvals))
}

models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m52 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

###################################################################################
##################################### Deep lake ################################
x00 <- glmmTMB(polyp_chao1~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Deep lake"))
## random effects variance estimated to be 0, logLik not calculated. Remove random factor
x00 <- glmmTMB(polyp_chao1~1, 
              data=div%>%filter(water_PopID=="Deep lake"))
x1 <- glmmTMB(polyp_chao1~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Deep lake"))
x2 <- glmmTMB(polyp_chao1~polyp_Season, 
              data=div%>%filter(water_PopID=="Deep lake")) 
x01 <- glmmTMB(water_chao1~1, 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)))
x3 <- glmmTMB(water_chao1~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)))
x4 <- glmmTMB(water_chao1~polyp_Season, 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)))
x5 <- glmmTMB(polyp_chao1~scale(water_chao1), 
              data=div%>%filter(water_PopID=="Deep lake"))

extract_info <- function(model,nullmodel) {
  pvals <- anova(model, nullmodel)[[8]][2]
  r2m <- r.squaredGLMM(model)[1]
  return(list(R2m = r2m, pvals = pvals))
}

models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m25 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

###################################################################################
##################################### River ################################
x00 <- glmmTMB(polyp_chao1~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"), family="gaussian")
x1 <- glmmTMB(polyp_chao1~scale(water_Temp)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"))
x2 <- glmmTMB(polyp_chao1~polyp_Season+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"))
x01 <- glmmTMB(water_chao1~1, 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)))
x3 <- glmmTMB(water_chao1~scale(water_Temp), 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)))
x4 <- glmmTMB(water_chao1~polyp_Season, 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)))
x5 <- glmmTMB(polyp_chao1~scale(water_chao1)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"), family="gaussian")

extract_info <- function(model,nullmodel) {
  pvals <- anova(model, nullmodel)[[8]][2]
  r2m <- r.squaredGLMM(model)[1]
  return(list(R2m = r2m, pvals = pvals))
}

models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m26 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

full_results_table <- rbind(results_table_m52, results_table_m25, results_table_m26)
full_results_table <- cbind(popID=rep(c("Shallow lake","Deep lake","River"),each=2),
                            response=rep(c("polyp_alpha","water_alpha"),3),
                            full_results_table)
names(full_results_table)[3:4] <- paste("Temperature",names(full_results_table)[3:4],
                                        sep="_")
names(full_results_table)[5:6] <- paste("Seasonality",names(full_results_table)[5:6],
                                        sep="_")

write.table(
  full_results_table,
  file="/storage/microbiome/seasonal_2024/figs/alpha_diversity_statistics.csv",
  col.names=T, row.names=T, quote=F, sep="\t"
)
  
```


```{r, fig.height=10, fig.width=10, message=F, warning=F}
library(glmmTMB); library(DHARMa); library(MuMIn)

###################################################################################
##################################### Shallow lake ################################
x00 <- glmmTMB(polyp_shannon~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"))
x1 <- glmmTMB(polyp_shannon~scale(water_Temp)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"))
x2 <- glmmTMB(polyp_shannon~polyp_Season+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"))
x01 <- glmmTMB(water_shannon~1, 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)))
x3 <- glmmTMB(water_shannon~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)))
x4 <- glmmTMB(water_shannon~polyp_Season, 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)))
x5 <- glmmTMB(polyp_shannon~scale(water_shannon)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"))

extract_info <- function(model,nullmodel) {
  pvals <- anova(model, nullmodel)[[8]][2]
  r2m <- r.squaredGLMM(model)[1]
  return(list(R2m = r2m, pvals = pvals))
}

models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m52 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

###################################################################################
##################################### Deep lake ################################
x00 <- glmmTMB(polyp_shannon~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Deep lake"))
## random effects variance estimated to be 0, logLik not calculated. Remove random factor
#x00 <- glmmTMB(polyp_shannon~1, 
#              data=div%>%filter(water_PopID=="Deep lake"))
x1 <- glmmTMB(polyp_shannon~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Deep lake"))
x2 <- glmmTMB(polyp_shannon~polyp_Season, 
              data=div%>%filter(water_PopID=="Deep lake")) 
x01 <- glmmTMB(water_shannon~1, 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)))
x3 <- glmmTMB(water_shannon~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)))
x4 <- glmmTMB(water_shannon~polyp_Season, 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)))
x5 <- glmmTMB(polyp_shannon~scale(water_shannon), 
              data=div%>%filter(water_PopID=="Deep lake"))

models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m25 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

###################################################################################
##################################### River ################################
x00 <- glmmTMB(polyp_shannon~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"), family="gaussian")
x1 <- glmmTMB(polyp_shannon~scale(water_Temp)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"))
x2 <- glmmTMB(polyp_shannon~polyp_Season+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"))
x01 <- glmmTMB(water_shannon~1, 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)))
x3 <- glmmTMB(water_shannon~scale(water_Temp), 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)))
x4 <- glmmTMB(water_shannon~polyp_Season, 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)))
x5 <- glmmTMB(polyp_shannon~scale(water_shannon)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"), family="gaussian")


models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m26 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

full_results_table <- rbind(results_table_m52, results_table_m25, results_table_m26)
full_results_table <- cbind(popID=rep(c("Shallow lake","Deep lake","River"),each=2),
                            response=rep(c("polyp_alpha","water_alpha"),3),
                            full_results_table)
names(full_results_table)[3:4] <- paste("Temperature",names(full_results_table)[3:4],
                                        sep="_")
names(full_results_table)[5:6] <- paste("Seasonality",names(full_results_table)[5:6],
                                        sep="_")

write.table(
  full_results_table,
  file="/storage/microbiome/seasonal_2024/figs/alpha_diversity_statistics_shannon.csv",
  col.names=T, row.names=T, quote=F, sep="\t"
)
  
```


```{r, fig.height=10, fig.width=10, message=F, warning=F}
library(glmmTMB); library(DHARMa); library(MuMIn)

###################################################################################
##################################### Shallow lake ################################
x00 <- glmmTMB(polyp_observed~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"), family="nbinom1")
x1 <- glmmTMB(polyp_observed~scale(water_Temp)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"), family="nbinom1")
x2 <- glmmTMB(polyp_observed~polyp_Season+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"), family="nbinom1")
x01 <- glmmTMB(water_observed~1, 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)), family="nbinom1")
x3 <- glmmTMB(water_observed~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)), family="nbinom1")
x4 <- glmmTMB(water_observed~polyp_Season, 
              data=div%>%filter(water_PopID=="Shallow lake"&!duplicated(water_SeqID)), family="nbinom1")
x5 <- glmmTMB(polyp_observed~scale(water_observed)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Shallow lake"), family="nbinom1")

models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m52 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

###################################################################################
##################################### Deep lake ################################
x00 <- glmmTMB(polyp_observed~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="Deep lake"), family="nbinom1")
x1 <- glmmTMB(polyp_observed~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Deep lake"), family="nbinom1")
x2 <- glmmTMB(polyp_observed~polyp_Season, 
              data=div%>%filter(water_PopID=="Deep lake"), family="nbinom1") 
x01 <- glmmTMB(water_observed~1, 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)), family="nbinom1")
x3 <- glmmTMB(water_observed~scale(water_Temp), 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)), family="nbinom1")
x4 <- glmmTMB(water_observed~polyp_Season, 
              data=div%>%filter(water_PopID=="Deep lake"&!duplicated(water_SeqID)), family="nbinom1")
x5 <- glmmTMB(polyp_observed~scale(water_observed), 
              data=div%>%filter(water_PopID=="Deep lake"), family="nbinom1")

models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m25 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

###################################################################################
##################################### River ################################
x00 <- glmmTMB(polyp_observed~1+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"), family="nbinom1")
x1 <- glmmTMB(polyp_observed~scale(water_Temp)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"), family="nbinom1")
x2 <- glmmTMB(polyp_observed~polyp_Season+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"), family="nbinom1")
x01 <- glmmTMB(water_observed~1, 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)), family="nbinom1")
x3 <- glmmTMB(water_observed~scale(water_Temp), 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)), family="nbinom1")
x4 <- glmmTMB(water_observed~polyp_Season, 
              data=div%>%filter(water_PopID=="River"&!duplicated(water_SeqID)), family="nbinom1")
x5 <- glmmTMB(polyp_observed~scale(water_observed)+(1|polyp_SiteID), 
              data=div%>%filter(water_PopID=="River"), family="nbinom1")

models <- list(x1, x2, x3, x4, x5)
null_models <- list(x00, x00, x01, x01, x00)

results <- mapply(extract_info, models, null_models, SIMPLIFY = FALSE)

results_table_m26 <- cbind(
  rbind(as.data.frame(results[[1]]), as.data.frame(results[[3]])),
  rbind(as.data.frame(results[[2]]), as.data.frame(results[[4]])),
  rbind(as.data.frame(results[[5]]), data.frame(R2m=NA, pvals=NA))
)

full_results_table <- rbind(results_table_m52, results_table_m25, results_table_m26)
full_results_table <- cbind(popID=rep(c("Shallow lake","Deep lake","River"),each=2),
                            response=rep(c("polyp_alpha","water_alpha"),3),
                            full_results_table)
names(full_results_table)[3:4] <- paste("Temperature",names(full_results_table)[3:4],
                                        sep="_")
names(full_results_table)[5:6] <- paste("Seasonality",names(full_results_table)[5:6],
                                        sep="_")

write.table(
  full_results_table,
  file="/storage/microbiome/seasonal_2024/figs/alpha_diversity_statistics_observed.csv",
  col.names=T, row.names=T, quote=F, sep="\t"
)
  
```


# Taxonomy barplots
```{r, fig.width=10, message=F, warning=F}
polyps <- subset_samples(ps_seasonal,Type=="polyp")
water <- subset_samples(ps_seasonal,Type=="water")

ragg::agg_tiff(filename="/storage/microbiome/seasonal_2024/figs/tax_polyps.tiff",
               width=10, height=8,units="in",res=300, scaling=0.75)
bar_polyps <- comp_barplot(polyps, tax_level="Genus",n_taxa=10)+
  ggh4x::facet_grid2(PopID~Sampling, scales="free_x", independent="x")+
  theme(axis.text.x = element_blank())+labs(title="Top bacterial taxa in polyp samples")

bar_water <- comp_barplot(water, tax_level="Genus",n_taxa=10, palette=distinct_palette(10,pal="kelly"))+
  ggh4x::facet_grid2(PopID~Sampling, scales="free_x", independent="x")+
  theme(axis.text.x = element_blank())+labs(title="Top bacterial taxa in water samples")

ggarrange(bar_polyps, bar_water, ncol=1,labels="AUTO")

dev.off()
```
```{r, fig.width=10, message=F, warning=F, eval=F}
subset_samples(ps_seasonal, Type=="polyp"&PopID=="M25")%>%
  tax_transform("clr",rank="Genus")%>%
  dist_calc(dist="bray")%>%
  ord_calc(method="PCA")%>%
  ord_plot(color="Sampling",plot_taxa=1:10, size=4,,
           tax_lab_style = tax_lab_style(size=4))+labs(title="Population M25")
subset_samples(ps_seasonal, Type=="polyp"&PopID=="M26")%>%
  tax_transform("clr",rank="Genus")%>%
  dist_calc(dist="bray")%>%
  ord_calc(method="PCA")%>%
  ord_plot(color="Sampling",plot_taxa=1:10, size=4,
           tax_lab_style = tax_lab_style(size=4))+labs(title="Population M26")
subset_samples(ps_seasonal, Type=="polyp"&PopID=="M52")%>%
  tax_transform("clr",rank="Genus")%>%
  dist_calc(dist="bray")%>%
  ord_calc(method="PCA")%>%
  ord_plot(color="Sampling",plot_taxa=1:10, size=4,
           tax_lab_style = tax_lab_style(size=4))+labs(title="Population M52")
```

# Evaluate seasonality of microbiome composition of polyp and water samples

```{r}
### Polyp samples
ad.M25.bray.polyps <- vegan::adonis2(dist.bray.M25.polyps~Season, data=data.frame(sample_data(polypsM25)))
ad.M26.bray.polyps <- vegan::adonis2(dist.bray.M26.polyps~Season, data=data.frame(sample_data(polypsM26)))
ad.M52.bray.polyps <- vegan::adonis2(dist.bray.M52.polyps~Season,
               data=data.frame(sample_data(polypsM52)))

ad.M25.jaccard.polyps <- vegan::adonis2(dist.jaccard.M25.polyps~Season, data=data.frame(sample_data(polypsM25)))
ad.M26.jaccard.polyps <- vegan::adonis2(dist.jaccard.M26.polyps~Season, data=data.frame(sample_data(polypsM26)))
ad.M52.jaccard.polyps <- vegan::adonis2(dist.jaccard.M52.polyps~Season,
               data=data.frame(sample_data(polypsM52)))

dbRDA.bray.M25.polyps <- capscale(dist.bray.M25.polyps ~ Temp, data=data.frame(sample_data(polypsM25)))
dbRDA.bray.M26.polyps <- capscale(dist.bray.M26.polyps ~ Temp, data=data.frame(sample_data(polypsM26)))
dbRDA.bray.M52.polyps <- capscale(dist.bray.M52.polyps ~ Temp, data=data.frame(sample_data(polypsM52)))

dbRDA.jaccard.M25.polyps <- capscale(dist.jaccard.M25.polyps ~ Temp, data=data.frame(sample_data(polypsM25)))
dbRDA.jaccard.M26.polyps <- capscale(dist.jaccard.M26.polyps ~ Temp, data=data.frame(sample_data(polypsM26)))
dbRDA.jaccard.M52.polyps <- capscale(dist.jaccard.M52.polyps ~ Temp, data=data.frame(sample_data(polypsM52)))

### Water samples
ad.M25.bray.water <- vegan::adonis2(dist.bray.M25.water~Season, data=data.frame(sample_data(waterM25)))
ad.M26.bray.water <- vegan::adonis2(dist.bray.M26.water~Season, data=data.frame(sample_data(waterM26)))
ad.M52.bray.water <- vegan::adonis2(dist.bray.M52.water~Season,
               data=data.frame(sample_data(waterM52)))

ad.M25.jaccard.water <- vegan::adonis2(dist.jaccard.M25.water~Season, data=data.frame(sample_data(waterM25)))
ad.M26.jaccard.water <- vegan::adonis2(dist.jaccard.M26.water~Season, data=data.frame(sample_data(waterM26)))
ad.M52.jaccard.water <- vegan::adonis2(dist.jaccard.M52.water~Season,
               data=data.frame(sample_data(waterM52)))

dbRDA.bray.M25.water <- capscale(dist.bray.M25.water ~ Temp, data=data.frame(sample_data(waterM25)))
dbRDA.bray.M26.water <- capscale(dist.bray.M26.water ~ Temp, data=data.frame(sample_data(waterM26)))
dbRDA.bray.M52.water <- capscale(dist.bray.M52.water ~ Temp, data=data.frame(sample_data(waterM52)))

dbRDA.jaccard.M25.water <- capscale(dist.jaccard.M25.water ~ Temp, data=data.frame(sample_data(waterM25)))
dbRDA.jaccard.M26.water <- capscale(dist.jaccard.M26.water ~ Temp, data=data.frame(sample_data(waterM26)))
dbRDA.jaccard.M52.water <- capscale(dist.jaccard.M52.water ~ Temp, data=data.frame(sample_data(waterM52)))

beta.results.bray <- data.frame(
  population=rep(c("Shallow lake", "Deep lake", "River"),each=2),
  type=rep(c("polyp","water"),3),
  season.r.squared=NA, season.p=NA,
  temp.r.squared=NA, temp.p=NA
)
beta.results.bray$season.r.squared[1] <- ad.M52.bray.polyps$R2[1]
beta.results.bray$season.p[1] <- ad.M52.bray.polyps$`Pr(>F)`[1]

beta.results.bray$season.r.squared[2] <- ad.M52.bray.water$R2[1]
beta.results.bray$season.p[2] <- ad.M52.bray.water$`Pr(>F)`[1]

beta.results.bray$season.r.squared[3] <- ad.M25.bray.polyps$R2[1]
beta.results.bray$season.p[3] <- ad.M25.bray.polyps$`Pr(>F)`[1]

beta.results.bray$season.r.squared[4] <- ad.M25.bray.water$R2[1]
beta.results.bray$season.p[4] <- ad.M25.bray.water$`Pr(>F)`[1]

beta.results.bray$season.r.squared[5] <- ad.M26.bray.polyps$R2[1]
beta.results.bray$season.p[5] <- ad.M26.bray.polyps$`Pr(>F)`[1]

beta.results.bray$season.r.squared[6] <- ad.M26.bray.water$R2[1]
beta.results.bray$season.p[6] <- ad.M26.bray.water$`Pr(>F)`[1]

beta.results.bray$temp.r.squared[1] <- vegan::RsquareAdj(dbRDA.bray.M52.polyps)$r.squared
beta.results.bray$temp.p[1] <- anova(dbRDA.bray.M52.polyps)$`Pr(>F)`[1]

beta.results.bray$temp.r.squared[2] <- vegan::RsquareAdj(dbRDA.bray.M52.water)$r.squared
beta.results.bray$temp.p[2] <- anova(dbRDA.bray.M52.water)$`Pr(>F)`[1]

beta.results.bray$temp.r.squared[3] <- vegan::RsquareAdj(dbRDA.bray.M25.polyps)$r.squared
beta.results.bray$temp.p[3] <- anova(dbRDA.bray.M25.polyps)$`Pr(>F)`[1]

beta.results.bray$temp.r.squared[4] <- vegan::RsquareAdj(dbRDA.bray.M25.water)$r.squared
beta.results.bray$temp.p[4] <- anova(dbRDA.bray.M25.water)$`Pr(>F)`[1]

beta.results.bray$temp.r.squared[5] <- vegan::RsquareAdj(dbRDA.bray.M26.polyps)$r.squared
beta.results.bray$temp.p[5] <- anova(dbRDA.bray.M26.polyps)$`Pr(>F)`[1]

beta.results.bray$temp.r.squared[6] <- vegan::RsquareAdj(dbRDA.bray.M26.water)$r.squared
beta.results.bray$temp.p[6] <- anova(dbRDA.bray.M26.water)$`Pr(>F)`[1]


#########

beta.results.jaccard <- data.frame(
  population=rep(c("Shallow lake", "Deep lake", "River"),each=2),
  type=rep(c("polyp","water"),3),
  season.r.squared=NA, season.p=NA,
  temp.r.squared=NA, temp.p=NA
)
beta.results.jaccard$season.r.squared[1] <- ad.M52.jaccard.polyps$R2[1]
beta.results.jaccard$season.p[1] <- ad.M52.jaccard.polyps$`Pr(>F)`[1]

beta.results.jaccard$season.r.squared[2] <- ad.M52.jaccard.water$R2[1]
beta.results.jaccard$season.p[2] <- ad.M52.jaccard.water$`Pr(>F)`[1]

beta.results.jaccard$season.r.squared[3] <- ad.M25.jaccard.polyps$R2[1]
beta.results.jaccard$season.p[3] <- ad.M25.jaccard.polyps$`Pr(>F)`[1]

beta.results.jaccard$season.r.squared[4] <- ad.M25.jaccard.water$R2[1]
beta.results.jaccard$season.p[4] <- ad.M25.jaccard.water$`Pr(>F)`[1]

beta.results.jaccard$season.r.squared[5] <- ad.M26.jaccard.polyps$R2[1]
beta.results.jaccard$season.p[5] <- ad.M26.jaccard.polyps$`Pr(>F)`[1]

beta.results.jaccard$season.r.squared[6] <- ad.M26.jaccard.water$R2[1]
beta.results.jaccard$season.p[6] <- ad.M26.jaccard.water$`Pr(>F)`[1]

beta.results.jaccard$temp.r.squared[1] <- vegan::RsquareAdj(dbRDA.jaccard.M52.polyps)$r.squared
beta.results.jaccard$temp.p[1] <- anova(dbRDA.jaccard.M52.polyps)$`Pr(>F)`[1]

beta.results.jaccard$temp.r.squared[2] <- vegan::RsquareAdj(dbRDA.jaccard.M52.water)$r.squared
beta.results.jaccard$temp.p[2] <- anova(dbRDA.jaccard.M52.water)$`Pr(>F)`[1]

beta.results.jaccard$temp.r.squared[3] <- vegan::RsquareAdj(dbRDA.jaccard.M25.polyps)$r.squared
beta.results.jaccard$temp.p[3] <- anova(dbRDA.jaccard.M25.polyps)$`Pr(>F)`[1]

beta.results.jaccard$temp.r.squared[4] <- vegan::RsquareAdj(dbRDA.jaccard.M25.water)$r.squared
beta.results.jaccard$temp.p[4] <- anova(dbRDA.jaccard.M25.water)$`Pr(>F)`[1]

beta.results.jaccard$temp.r.squared[5] <- vegan::RsquareAdj(dbRDA.jaccard.M26.polyps)$r.squared
beta.results.jaccard$temp.p[5] <- anova(dbRDA.jaccard.M26.polyps)$`Pr(>F)`[1]

beta.results.jaccard$temp.r.squared[6] <- vegan::RsquareAdj(dbRDA.jaccard.M26.water)$r.squared
beta.results.jaccard$temp.p[6] <- anova(dbRDA.jaccard.M26.water)$`Pr(>F)`[1]
```

# Mantel test - Bray
```{r Mantel}
temp.dat <- sample_data(subset_samples(ps_seasonal, Type=="polyp"))
temp.dat <- temp.dat[which(!duplicated(gsub("[0-9]*$","",temp.dat$SeqID))),]

season <- temp.dat$Season
names(season) <- row.names(temp.dat)
season_diss <- as.matrix(outer(season, season, FUN = function(x, y) as.integer(x != y)))

temperature <- temp.dat$Temp
names(temperature) <- row.names(temp.dat)
temperature_diss <- as.matrix(as.dist(abs(outer(temperature, temperature, FUN = "-"))))

tt.m25 <- temperature_diss[grep("M25",row.names(temperature_diss)),grep("M25",colnames(temperature_diss))]
tt.m26 <- temperature_diss[grep("M26",row.names(temperature_diss)),grep("M26",colnames(temperature_diss))]
tt.m52 <- temperature_diss[grep("M52",row.names(temperature_diss)),grep("M52",colnames(temperature_diss))]

ss.m25 <- season_diss[grep("M25",row.names(season_diss)),grep("M25",colnames(season_diss))]
ss.m26 <- season_diss[grep("M26",row.names(season_diss)),grep("M26",colnames(season_diss))]
ss.m52 <- season_diss[grep("M52",row.names(season_diss)),grep("M52",colnames(season_diss))]

w.data <- sample_data(subset_samples(ps_seasonal, Type=="water"))
w.mat <- as.matrix(distance(subset_samples(ps_seasonal, Type=="water"),"bray"))
row.names(w.mat) <- gsub("W","",row.names(w.mat))
colnames(w.mat) <- gsub("W","",colnames(w.mat))

p.data <- sample_data(subset_samples(ps_seasonal, Type=="polyp"))
p.mat <- as.matrix(distance(subset_samples(ps_seasonal, Type=="polyp"), "bray"))

p.m25 <- p.mat[grep("M25",row.names(p.mat)),grep("M25",colnames(p.mat))]
p.m26 <- p.mat[grep("M26",row.names(p.mat)),grep("M26",colnames(p.mat))]
p.m52 <- p.mat[grep("M52",row.names(p.mat)),grep("M52",colnames(p.mat))]

sites <- unique(gsub("[0-9]*$","",row.names(p.mat)))

pp.mat <- matrix(NA, ncol=length(sites),nrow=length(sites), dimnames=list(sites,sites))

for(i in 1:nrow(pp.mat)){
  for(j in 1:ncol(pp.mat)){
    row <- row.names(pp.mat)[i]
    col <- colnames(pp.mat)[j]
    site.mat <- p.mat[which(gsub("[0-9]*$","",row.names(p.mat))==row),which(gsub("[0-9]*$","",colnames(p.mat))==col),drop=FALSE]
    if(identical(row, col)){
      val<-mean(site.mat[upper.tri(site.mat)])
    } else {
      val<-mean(site.mat)
    }
    pp.mat[,which(colnames(pp.mat)==col)][which(row.names(pp.mat)==row)] <- val
  }
}

pp.m25 <- pp.mat[grep("M25",row.names(pp.mat)),grep("M25",colnames(pp.mat))]
pp.m26 <- pp.mat[grep("M26",row.names(pp.mat)),grep("M26",colnames(pp.mat))]
pp.m52 <- pp.mat[grep("M52",row.names(pp.mat)),grep("M52",colnames(pp.mat))]

ww.mat <- matrix(NA, ncol=length(sites),nrow=length(sites), dimnames=list(sites,sites))

for(i in 1:nrow(pp.mat)){
  for(j in 1:ncol(pp.mat)){
    row <- row.names(ww.mat)[i]
    col <- colnames(ww.mat)[j]
  if(row%in%row.names(w.mat) & col%in%colnames(w.mat)){
    ww.mat[,which(colnames(ww.mat)==col)][which(row.names(ww.mat)==row)] <- 
      w.mat[which(row.names(w.mat)==row), which(colnames(w.mat)==col)]
    }
  }
}

ww.m25 <- ww.mat[grep("M25",row.names(ww.mat)),grep("M25",colnames(ww.mat))]
ww.m26 <- ww.mat[grep("M26",row.names(ww.mat)),grep("M26",colnames(ww.mat))]
ww.m52 <- ww.mat[grep("M52",row.names(ww.mat)),grep("M52",colnames(ww.mat))]

vegan::mantel(pp.m25, ww.m25, na.rm=T)
vegan::mantel(pp.m26, ww.m26, na.rm=T)
vegan::mantel(pp.m52, ww.m52, na.rm=T)

pmant.temp.bray.m25 <- vegan::mantel.partial(pp.m25, ww.m25, tt.m25, na.rm=T)
pmant.temp.bray.m26 <- vegan::mantel.partial(pp.m26, ww.m26, tt.m26, na.rm=T)
pmant.temp.bray.m52 <- vegan::mantel.partial(pp.m52, ww.m52, tt.m52, na.rm=T)

pmant.season.bray.m25 <- vegan::mantel.partial(pp.m25, ww.m25, ss.m25, na.rm=T)
pmant.season.bray.m26 <- vegan::mantel.partial(pp.m26, ww.m26, ss.m26, na.rm=T)
pmant.season.bray.m52 <- vegan::mantel.partial(pp.m52, ww.m52, ss.m52, na.rm=T)

```

# Mantel test - Jaccard
```{r Mantel}
temp.dat <- sample_data(subset_samples(ps_seasonal, Type=="polyp"))
temp.dat <- temp.dat[which(!duplicated(gsub("[0-9]*$","",temp.dat$SeqID))),]

season <- temp.dat$Season
names(season) <- row.names(temp.dat)
season_diss <- as.matrix(outer(season, season, FUN = function(x, y) as.integer(x != y)))

temperature <- temp.dat$Temp
names(temperature) <- row.names(temp.dat)
temperature_diss <- as.matrix(as.dist(abs(outer(temperature, temperature, FUN = "-"))))

tt.m25 <- temperature_diss[grep("M25",row.names(temperature_diss)),grep("M25",colnames(temperature_diss))]
tt.m26 <- temperature_diss[grep("M26",row.names(temperature_diss)),grep("M26",colnames(temperature_diss))]
tt.m52 <- temperature_diss[grep("M52",row.names(temperature_diss)),grep("M52",colnames(temperature_diss))]

ss.m25 <- season_diss[grep("M25",row.names(season_diss)),grep("M25",colnames(season_diss))]
ss.m26 <- season_diss[grep("M26",row.names(season_diss)),grep("M26",colnames(season_diss))]
ss.m52 <- season_diss[grep("M52",row.names(season_diss)),grep("M52",colnames(season_diss))]

w.data <- sample_data(subset_samples(ps_seasonal, Type=="water"))
w.mat <- as.matrix(distance(subset_samples(ps_seasonal, Type=="water"),"jaccard", binary=T))
row.names(w.mat) <- gsub("W","",row.names(w.mat))
colnames(w.mat) <- gsub("W","",colnames(w.mat))

p.data <- sample_data(subset_samples(ps_seasonal, Type=="polyp"))
p.mat <- as.matrix(distance(subset_samples(ps_seasonal, Type=="polyp"), "jaccard", binary=T))

p.m25 <- p.mat[grep("M25",row.names(p.mat)),grep("M25",colnames(p.mat))]
p.m26 <- p.mat[grep("M26",row.names(p.mat)),grep("M26",colnames(p.mat))]
p.m52 <- p.mat[grep("M52",row.names(p.mat)),grep("M52",colnames(p.mat))]

sites <- unique(gsub("[0-9]*$","",row.names(p.mat)))

pp.mat <- matrix(NA, ncol=length(sites),nrow=length(sites), dimnames=list(sites,sites))

for(i in 1:nrow(pp.mat)){
  for(j in 1:ncol(pp.mat)){
    row <- row.names(pp.mat)[i]
    col <- colnames(pp.mat)[j]
    site.mat <- p.mat[which(gsub("[0-9]*$","",row.names(p.mat))==row),which(gsub("[0-9]*$","",colnames(p.mat))==col),drop=FALSE]
    if(identical(row, col)){
      val<-mean(site.mat[upper.tri(site.mat)])
    } else {
      val<-mean(site.mat)
    }
    pp.mat[,which(colnames(pp.mat)==col)][which(row.names(pp.mat)==row)] <- val
  }
}

pp.m25 <- pp.mat[grep("M25",row.names(pp.mat)),grep("M25",colnames(pp.mat))]
pp.m26 <- pp.mat[grep("M26",row.names(pp.mat)),grep("M26",colnames(pp.mat))]
pp.m52 <- pp.mat[grep("M52",row.names(pp.mat)),grep("M52",colnames(pp.mat))]

ww.mat <- matrix(NA, ncol=length(sites),nrow=length(sites), dimnames=list(sites,sites))

for(i in 1:nrow(pp.mat)){
  for(j in 1:ncol(pp.mat)){
    row <- row.names(ww.mat)[i]
    col <- colnames(ww.mat)[j]
  if(row%in%row.names(w.mat) & col%in%colnames(w.mat)){
    ww.mat[,which(colnames(ww.mat)==col)][which(row.names(ww.mat)==row)] <- 
      w.mat[which(row.names(w.mat)==row), which(colnames(w.mat)==col)]
    }
  }
}

ww.m25 <- ww.mat[grep("M25",row.names(ww.mat)),grep("M25",colnames(ww.mat))]
ww.m26 <- ww.mat[grep("M26",row.names(ww.mat)),grep("M26",colnames(ww.mat))]
ww.m52 <- ww.mat[grep("M52",row.names(ww.mat)),grep("M52",colnames(ww.mat))]

vegan::mantel(pp.m25, ww.m25, na.rm=T)
vegan::mantel(pp.m26, ww.m26, na.rm=T)
vegan::mantel(pp.m52, ww.m52, na.rm=T)

pmant.temp.jaccard.m25 <- vegan::mantel.partial(pp.m25, ww.m25, tt.m25, na.rm=T)
pmant.temp.jaccard.m26 <- vegan::mantel.partial(pp.m26, ww.m26, tt.m26, na.rm=T)
pmant.temp.jaccard.m52 <- vegan::mantel.partial(pp.m52, ww.m52, tt.m52, na.rm=T)

pmant.season.jaccard.m25 <- vegan::mantel.partial(pp.m25, ww.m25, ss.m25, na.rm=T)
pmant.season.jaccard.m26 <- vegan::mantel.partial(pp.m26, ww.m26, ss.m26, na.rm=T)
pmant.season.jaccard.m52 <- vegan::mantel.partial(pp.m52, ww.m52, ss.m52, na.rm=T)
```
```{r}
partial.mantel.res.bray <- data.frame(
  pop=c("Shallow lake","Deep lake", "River"),
  season.stat=NA, season.p=NA,
  temp.stat=NA, temp.p=NA)
partial.mantel.res.bray$season.stat[1] <- pmant.season.bray.m52$statistic
partial.mantel.res.bray$season.stat[2] <- pmant.season.bray.m25$statistic
partial.mantel.res.bray$season.stat[3] <- pmant.season.bray.m26$statistic

partial.mantel.res.bray$season.p[1] <- pmant.season.bray.m52$signif
partial.mantel.res.bray$season.p[2] <- pmant.season.bray.m25$signif
partial.mantel.res.bray$season.p[3] <- pmant.season.bray.m26$signif

partial.mantel.res.bray$temp.stat[1] <- pmant.temp.bray.m52$statistic
partial.mantel.res.bray$temp.stat[2] <- pmant.temp.bray.m25$statistic
partial.mantel.res.bray$temp.stat[3] <- pmant.temp.bray.m26$statistic

partial.mantel.res.bray$temp.p[1] <- pmant.temp.bray.m52$signif
partial.mantel.res.bray$temp.p[2] <- pmant.temp.bray.m25$signif
partial.mantel.res.bray$temp.p[3] <- pmant.temp.bray.m26$signif

partial.mantel.res.jaccard <- data.frame(
  pop=c("Shallow lake","Deep lake", "River"),
  season.stat=NA, season.p=NA,
  temp.stat=NA, temp.p=NA)
partial.mantel.res.jaccard$season.stat[1] <- pmant.season.jaccard.m52$statistic
partial.mantel.res.jaccard$season.stat[2] <- pmant.season.jaccard.m25$statistic
partial.mantel.res.jaccard$season.stat[3] <- pmant.season.jaccard.m26$statistic

partial.mantel.res.jaccard$season.p[1] <- pmant.season.jaccard.m52$signif
partial.mantel.res.jaccard$season.p[2] <- pmant.season.jaccard.m25$signif
partial.mantel.res.jaccard$season.p[3] <- pmant.season.jaccard.m26$signif

partial.mantel.res.jaccard$temp.stat[1] <- pmant.temp.jaccard.m52$statistic
partial.mantel.res.jaccard$temp.stat[2] <- pmant.temp.jaccard.m25$statistic
partial.mantel.res.jaccard$temp.stat[3] <- pmant.temp.jaccard.m26$statistic

partial.mantel.res.jaccard$temp.p[1] <- pmant.temp.jaccard.m52$signif
partial.mantel.res.jaccard$temp.p[2] <- pmant.temp.jaccard.m25$signif
partial.mantel.res.jaccard$temp.p[3] <- pmant.temp.jaccard.m26$signif
```


# Compare distance between water and polyps samples across sampling events - BRAY-CURTIS
```{r}
library(tidyr); library(tibble)
distmat.bray.M25 <- dist.bray.M25 %>% as.matrix()
distmat.bray.M25[upper.tri(distmat.bray.M25, diag = FALSE)] <- NA
df_distmat.bray.M25 <- as.data.frame(distmat.bray.M25) %>%
  rownames_to_column(var = "Row") %>%
  pivot_longer(-Row, names_to = "Column", values_to = "Value") %>%
  filter(!is.na(Value))
df_distmat.bray.M25$source_type <- ifelse(grepl("W",df_distmat.bray.M25$Row), "w","p")
df_distmat.bray.M25$source_site <- substr(gsub("W","",df_distmat.bray.M25$Row),5,5)
df_distmat.bray.M25$source_sampling <- substr(gsub("W","",df_distmat.bray.M25$Row),6,6)
df_distmat.bray.M25$target_type <- ifelse(grepl("W",df_distmat.bray.M25$Column), "w","p")
df_distmat.bray.M25$target_site <- substr(gsub("W","",df_distmat.bray.M25$Column),5,5)
df_distmat.bray.M25$target_sampling <- substr(gsub("W","",df_distmat.bray.M25$Column),6,6)
pwdist.M25 <- df_distmat.bray.M25%>%filter(source_sampling==target_sampling & source_site==target_site & source_type != target_type)
pwdist.M25$Pop <- "Deep lake"

distmat.bray.M26 <- dist.bray.M26 %>% as.matrix()
distmat.bray.M26[upper.tri(distmat.bray.M26, diag = FALSE)] <- NA
df_distmat.bray.M26 <- as.data.frame(distmat.bray.M26) %>%
  rownames_to_column(var = "Row") %>%
  pivot_longer(-Row, names_to = "Column", values_to = "Value") %>%
  filter(!is.na(Value))
df_distmat.bray.M26$source_type <- ifelse(grepl("W",df_distmat.bray.M26$Row), "w","p")
df_distmat.bray.M26$source_site <- substr(gsub("W","",df_distmat.bray.M26$Row),5,5)
df_distmat.bray.M26$source_sampling <- substr(gsub("W","",df_distmat.bray.M26$Row),6,6)
df_distmat.bray.M26$target_type <- ifelse(grepl("W",df_distmat.bray.M26$Column), "w","p")
df_distmat.bray.M26$target_site <- substr(gsub("W","",df_distmat.bray.M26$Column),5,5)
df_distmat.bray.M26$target_sampling <- substr(gsub("W","",df_distmat.bray.M26$Column),6,6)
pwdist.M26 <- df_distmat.bray.M26%>%filter(source_sampling==target_sampling & source_site==target_site & source_type != target_type)
boxplot(pwdist.M26$Value~pwdist.M26$source_sampling)
pwdist.M26$Pop <- "River"

distmat.bray.M52 <- dist.bray.M52 %>% as.matrix()
distmat.bray.M52[upper.tri(distmat.bray.M52, diag = FALSE)] <- NA
df_distmat.bray.M52 <- as.data.frame(distmat.bray.M52) %>%
  rownames_to_column(var = "Row") %>%
  pivot_longer(-Row, names_to = "Column", values_to = "Value") %>%
  filter(!is.na(Value))
df_distmat.bray.M52$source_type <- ifelse(grepl("W",df_distmat.bray.M52$Row), "w","p")
df_distmat.bray.M52$source_site <- substr(gsub("W","",df_distmat.bray.M52$Row),5,5)
df_distmat.bray.M52$source_sampling <- substr(gsub("W","",df_distmat.bray.M52$Row),6,6)
df_distmat.bray.M52$target_type <- ifelse(grepl("W",df_distmat.bray.M52$Column), "w","p")
df_distmat.bray.M52$target_site <- substr(gsub("W","",df_distmat.bray.M52$Column),5,5)
df_distmat.bray.M52$target_sampling <- substr(gsub("W","",df_distmat.bray.M52$Column),6,6)
pwdist.M52 <- df_distmat.bray.M52%>%filter(source_sampling==target_sampling & source_site==target_site & source_type != target_type)
pwdist.M52$Pop <- "Shallow lake"

pwdist.bray <- rbind(pwdist.M25, pwdist.M26, pwdist.M52)
pwdist.bray$Pop <- factor(pwdist.bray$Pop, levels=c("Shallow lake", "Deep lake", "River"))
pwdist.bray$Season <- pwdist.bray$source_sampling
pwdist.bray$Season[pwdist.bray$Season%in%c("B","F")] <- "Autumn"
pwdist.bray$Season[pwdist.bray$Season%in%c("C","G")] <- "Winter"
pwdist.bray$Season[pwdist.bray$Season%in%c("D","H")] <- "Spring"
pwdist.bray$Season[pwdist.bray$Season%in%c("E","J")] <- "Summer"

p1 <- ggplot(pwdist.bray, aes(x=source_sampling, y=Value, fill=Season))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter()+
  facet_wrap(~Pop, ncol=3)+  theme_bw()+
  ylab("Distance")+xlab("Sampling")+
  labs(title="Bray-Curtis dissimilarity of polyp and water samples")+
  scale_x_discrete(labels=c("2021 September","2021 December","2022 March",
                            "2022 June","2022 November","2023 February",
                            "2023 April","2023 June"))+
  theme(axis.text.x = element_text(angle=45,hjust=1),
        legend.position="bottom",
        legend.background = element_rect(color = "black", linewidth = 0.25),
        plot.title = element_text(hjust = 0.5,face="bold"))+
  scale_fill_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))

```

# Compare distance between water and polyps samples across sampling events - JACCARD
```{r}
library(tidyr); library(tibble)
distmat.jaccard.M25 <- dist.jaccard.M25 %>% as.matrix()
distmat.jaccard.M25[upper.tri(distmat.jaccard.M25, diag = FALSE)] <- NA
df_distmat.jaccard.M25 <- as.data.frame(distmat.jaccard.M25) %>%
  rownames_to_column(var = "Row") %>%
  pivot_longer(-Row, names_to = "Column", values_to = "Value") %>%
  filter(!is.na(Value))
df_distmat.jaccard.M25$source_type <- ifelse(grepl("W",df_distmat.jaccard.M25$Row), "w","p")
df_distmat.jaccard.M25$source_site <- substr(gsub("W","",df_distmat.jaccard.M25$Row),5,5)
df_distmat.jaccard.M25$source_sampling <- substr(gsub("W","",df_distmat.jaccard.M25$Row),6,6)
df_distmat.jaccard.M25$target_type <- ifelse(grepl("W",df_distmat.jaccard.M25$Column), "w","p")
df_distmat.jaccard.M25$target_site <- substr(gsub("W","",df_distmat.jaccard.M25$Column),5,5)
df_distmat.jaccard.M25$target_sampling <- substr(gsub("W","",df_distmat.jaccard.M25$Column),6,6)
pwdist.M25 <- df_distmat.jaccard.M25%>%filter(source_sampling==target_sampling & source_site==target_site & source_type != target_type)
pwdist.M25$Pop <- "Deep lake"

distmat.jaccard.M26 <- dist.jaccard.M26 %>% as.matrix()
distmat.jaccard.M26[upper.tri(distmat.jaccard.M26, diag = FALSE)] <- NA
df_distmat.jaccard.M26 <- as.data.frame(distmat.jaccard.M26) %>%
  rownames_to_column(var = "Row") %>%
  pivot_longer(-Row, names_to = "Column", values_to = "Value") %>%
  filter(!is.na(Value))
df_distmat.jaccard.M26$source_type <- ifelse(grepl("W",df_distmat.jaccard.M26$Row), "w","p")
df_distmat.jaccard.M26$source_site <- substr(gsub("W","",df_distmat.jaccard.M26$Row),5,5)
df_distmat.jaccard.M26$source_sampling <- substr(gsub("W","",df_distmat.jaccard.M26$Row),6,6)
df_distmat.jaccard.M26$target_type <- ifelse(grepl("W",df_distmat.jaccard.M26$Column), "w","p")
df_distmat.jaccard.M26$target_site <- substr(gsub("W","",df_distmat.jaccard.M26$Column),5,5)
df_distmat.jaccard.M26$target_sampling <- substr(gsub("W","",df_distmat.jaccard.M26$Column),6,6)
pwdist.M26 <- df_distmat.jaccard.M26%>%filter(source_sampling==target_sampling & source_site==target_site & source_type != target_type)
boxplot(pwdist.M26$Value~pwdist.M26$source_sampling)
pwdist.M26$Pop <- "River"

distmat.jaccard.M52 <- dist.jaccard.M52 %>% as.matrix()
distmat.jaccard.M52[upper.tri(distmat.jaccard.M52, diag = FALSE)] <- NA
df_distmat.jaccard.M52 <- as.data.frame(distmat.jaccard.M52) %>%
  rownames_to_column(var = "Row") %>%
  pivot_longer(-Row, names_to = "Column", values_to = "Value") %>%
  filter(!is.na(Value))
df_distmat.jaccard.M52$source_type <- ifelse(grepl("W",df_distmat.jaccard.M52$Row), "w","p")
df_distmat.jaccard.M52$source_site <- substr(gsub("W","",df_distmat.jaccard.M52$Row),5,5)
df_distmat.jaccard.M52$source_sampling <- substr(gsub("W","",df_distmat.jaccard.M52$Row),6,6)
df_distmat.jaccard.M52$target_type <- ifelse(grepl("W",df_distmat.jaccard.M52$Column), "w","p")
df_distmat.jaccard.M52$target_site <- substr(gsub("W","",df_distmat.jaccard.M52$Column),5,5)
df_distmat.jaccard.M52$target_sampling <- substr(gsub("W","",df_distmat.jaccard.M52$Column),6,6)
pwdist.M52 <- df_distmat.jaccard.M52%>%filter(source_sampling==target_sampling & source_site==target_site & source_type != target_type)
pwdist.M52$Pop <- "Shallow lake"

pwdist.jaccard <- rbind(pwdist.M25, pwdist.M26, pwdist.M52)
pwdist.jaccard$Pop <- factor(pwdist.jaccard$Pop, levels=c("Shallow lake", "Deep lake", "River"))
pwdist.jaccard$Season <- pwdist.jaccard$source_sampling
pwdist.jaccard$Season[pwdist.jaccard$Season%in%c("B","F")] <- "Autumn"
pwdist.jaccard$Season[pwdist.jaccard$Season%in%c("C","G")] <- "Winter"
pwdist.jaccard$Season[pwdist.jaccard$Season%in%c("D","H")] <- "Spring"
pwdist.jaccard$Season[pwdist.jaccard$Season%in%c("E","J")] <- "Summer"

p2 <- ggplot(pwdist.jaccard, aes(x=source_sampling, y=Value, fill=Season))+
  geom_boxplot(outlier.shape = NA)+
  geom_jitter()+
  facet_wrap(~Pop, ncol=3)+  theme_bw()+
  ylab("Distance")+xlab("Sampling")+
  labs(title="Jaccard presence/absence dissimilarity of polyp and water samples")+
  scale_x_discrete(labels=c("2021 September","2021 December","2022 March",
                            "2022 June","2022 November","2023 February",
                            "2023 April","2023 June"))+
  theme(axis.text.x = element_text(angle=45,hjust=1),
        legend.position="top",
        legend.background = element_rect(color = "black", linewidth = 0.25))+
  scale_fill_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))

```


# PERMANOVA
```{r, fig.width=10, message=F, warning=F}
pwdist.bray$Temp.SampleID <- sample_data(ps_seasonal)$SeqID[match(pwdist.bray$Column, row.names(sample_data(ps_seasonal)))]
identical(gsub(".","_",pwdist.bray$Column,fixed=T),pwdist.bray$Temp.SampleID)
pwdist.bray$Temp <- sample_data(ps_seasonal)$Temp[match(pwdist.bray$Column, row.names(sample_data(ps_seasonal)))]
pwdist.bray$Season <- factor(pwdist.bray$Season, levels=c("Spring","Summer","Autumn","Winter"))

pwdist.jaccard$Temp.SampleID <- sample_data(ps_seasonal)$SeqID[match(pwdist.jaccard$Column, row.names(sample_data(ps_seasonal)))]
identical(gsub(".","_",pwdist.jaccard$Column,fixed=T),pwdist.jaccard$Temp.SampleID)
pwdist.jaccard$Temp <- sample_data(ps_seasonal)$Temp[match(pwdist.jaccard$Column, row.names(sample_data(ps_seasonal)))]
pwdist.jaccard$Season <- factor(pwdist.jaccard$Season, levels=c("Spring","Summer","Autumn","Winter"))

m1.bray.season.m52 <- glmmTMB(Value~Season+(1|Column), data=pwdist.bray%>%filter(Pop=="Shallow lake"))
m1.bray.season.m25 <- glmmTMB(Value~Season+(1|Column), data=pwdist.bray%>%filter(Pop=="Deep lake"))
m1.bray.season.m26 <- glmmTMB(Value~Season+(1|Column), data=pwdist.bray%>%filter(Pop=="River"))

m1.bray.temp.m52 <- glmmTMB(Value~Temp+(1|Column), data=pwdist.bray%>%filter(Pop=="Shallow lake"))
m1.bray.temp.m25 <- glmmTMB(Value~Temp+(1|Column), data=pwdist.bray%>%filter(Pop=="Deep lake"))
m1.bray.temp.m26 <- glmmTMB(Value~Temp+(1|Column), data=pwdist.bray%>%filter(Pop=="River"))

m1.jaccard.season.m52 <- glmmTMB(Value~Season+(1|Column), data=pwdist.jaccard%>%filter(Pop=="Shallow lake"))
m1.jaccard.season.m25 <- glmmTMB(Value~Season+(1|Column), data=pwdist.jaccard%>%filter(Pop=="Deep lake"))
m1.jaccard.season.m26 <- glmmTMB(Value~Season+(1|Column), data=pwdist.jaccard%>%filter(Pop=="River"))

m1.jaccard.temp.m52 <- glmmTMB(Value~Temp+(1|Column), data=pwdist.jaccard%>%filter(Pop=="Shallow lake"))
m1.jaccard.temp.m25 <- glmmTMB(Value~Temp+(1|Column), data=pwdist.jaccard%>%filter(Pop=="Deep lake"))
m1.jaccard.temp.m26 <- glmmTMB(Value~Temp+(1|Column), data=pwdist.jaccard%>%filter(Pop=="River"))

d1 <- ggplot(pwdist.bray, aes(x=Temp, y=Value,fill=Season))+
  geom_point(shape=21, colour="grey30",size=3)+
  facet_wrap(~Pop)+labs(title="Bray-Curtis index")+
  ylab("Dissimilarity")+xlab("Temperature (°C)")+theme_bw()+
  scale_fill_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))
d2 <- ggplot(pwdist.jaccard, aes(x=Temp, y=Value, fill=Season))+
  geom_point(shape=21, colour="grey30",size=3)+
  facet_wrap(~Pop)+labs(title="Jaccard presence-absence index")+
  ylab("Dissimilarity")+xlab("Temperature (°C)")+theme_bw()+
  scale_fill_manual(values=c("Spring"="lightgreen","Summer"="tan2",
                              "Autumn"="tomato2","Winter"="steelblue3"))

ragg::agg_tiff(filename="/storage/microbiome/seasonal_2024/figs/pw_diss.tiff",
               width=8, height=8,units="in",res=300, scaling=0.9)
ggarrange(d1,d2,ncol=1,labels="AUTO",common.legend=T)
dev.off()
```

#Upset plot

```{r}
library(MicrobiotaProcess); library(ComplexUpset); library(UpSetR); library(ComplexHeatmap)

sample_groups <- c("2021 September polyp","2021 September water",
                  "2021 December polyp","2021 December water",
                  "2022 March polyp","2022 March water",
                  "2022 June polyp","2022 June water",
                  "2022 November polyp","2022 November water",
                  "2023 February polyp","2023 February water",
                  "2023 April polyp","2023 April water",
                  "2023 June polyp","2023 June water")

sample_data(ps_seasonal)$SamplingType <- paste(sample_data(ps_seasonal)$Sampling,
                                               sample_data(ps_seasonal)$Type, sep=" ")
upset_datM52 <- get_upset(subset_samples(ps_seasonal, PopID=="Shallow lake"),
          factorNames="SamplingType",threshold=2)%>%select(!"2021 June polyp")

m <- make_comb_mat(upset_datM52,mode="intersect")
m2 <- m[,which(comb_name(m)%in%
c("1100000000000000","0011000000000000","0000110000000000","0000001100000000",
"0000000011000000","0000000000110000","0000000000001100","0000000000000011"))]

set.order <- match(names(set_size(m2)), sample_groups)
comb.order <- order(comb_degree(m2), -comb_size(m2))

ht_opt$TITLE_PADDING = unit(c(8.5, 8.5), "points")
u1 <- UpSet(m2,
      comb_col=c("steelblue3","tomato2","tan2","lightgreen",
                   "tomato2","lightgreen","steelblue3","tan2"),
      comb_order = comb.order,
      set_order=set.order, row_names_gp=gpar(fontsize=8),
      top_annotation = upset_top_annotation(m2, 
      annotation_name_rot = 90,
      annotation_name_offset = unit(10,"mm"),
      axis_param = list(gp = gpar(fontsize = 10)), 
      gp = gpar(col = "black",
                fill=c("steelblue3","tomato2","tan2","lightgreen",
                   "tomato2","lightgreen","steelblue3","tan2"))),
      column_title="Shallow lake",
      column_title_gp = gpar(fill = "grey85", col = "black", border = "black")
      )

## M25
upset_datM25 <- get_upset(subset_samples(ps_seasonal, PopID=="Deep lake"),
          factorNames="SamplingType",threshold=2)%>%select(!"2021 June polyp")

m <- make_comb_mat(upset_datM25,mode="intersect")
m2 <- m[,which(comb_name(m)%in%
c("1100000000000000","0011000000000000","0000110000000000","0000001100000000",
"0000000011000000","0000000000110000","0000000000001100","0000000000000011"))]

set.order <- match(names(set_size(m2)), sample_groups)
comb.order <- order(comb_degree(m2), -comb_size(m2))

u2 <- UpSet(m2,
      comb_col=c("steelblue3","tomato2","tan2","lightgreen",
                   "tomato2","lightgreen","steelblue3","tan2"),
      comb_order = comb.order,
      set_order=set.order,
      top_annotation = upset_top_annotation(m2, 
      annotation_name_rot = 90,
      annotation_name_offset = unit(10,"mm"),
      axis_param = list(gp = gpar(fontsize = 10)), 
      gp = gpar(col = "black",
                fill=c("steelblue3","tomato2","tan2","lightgreen",
                   "tomato2","lightgreen","steelblue3","tan2"))),
      column_title="Deep lake",
      column_title_gp = gpar(fill = "grey85", col = "black", border = "black")
      ) 

## M26
upset_datM26 <- get_upset(subset_samples(ps_seasonal, PopID=="River"),
          factorNames="SamplingType",threshold=2)%>%
  select(!"2021 June polyp")
upset_datM26$"2021 December water" <- 0
upset_datM26$"2021 September water" <- 0
upset_datM26$"2023 February water" <- 0

upset_datM26 <- upset_datM26[,c(1,14,2,15,3:10,11,16,12:13)]
         #&!"2021 December polyp"&!"2021 September polyp"&!"2023 February polyp")

m <- make_comb_mat(upset_datM26,mode="intersect")
m2 <- m[,which(comb_name(m)%in%
c("1100000000000000","0011000000000000","0000110000000000","0000001100000000",
"0000000011000000","0000000000110000","0000000000001100","0000000000000011"))]

set.order <- match(names(set_size(m2)), sample_groups)
comb.order <- order(comb_degree(m2), -comb_size(m2))

u3 <- UpSet(m2,
      comb_col=c("steelblue3","tomato2","tan2","lightgreen",
                   "tomato2","lightgreen","steelblue3","tan2"),
      comb_order = comb.order,
      set_order=set.order,
      top_annotation = upset_top_annotation(m2, 
      annotation_name_rot = 90,
      annotation_name_offset = unit(10,"mm"),
      axis_param = list(gp = gpar(fontsize = 10)),
      gp = gpar(col = "black",
                fill=c("steelblue3","tomato2","tan2","lightgreen",
                   "tomato2","lightgreen","steelblue3","tan2"))),
      column_title="River",
      column_title_gp = gpar(fill = "grey85", col = "black", border = "black")
      )
ht_grob <- grid::grid.grabExpr(
  draw(u1+u2+u3, column_title="No. of co-occurring bacterial taxa in polyp and water samples",
       column_title_gp=gpar(font=2),
       padding = unit(c(2, 5, 2, 2), "mm"))
)
ragg::agg_tiff(filename="/storage/microbiome/seasonal_2024/figs/pw.tiff",
               width=10, height=8,units="in",res=300, scaling=0.8)
ggarrange(ht_grob,p1,ncol=1,labels="AUTO")
dev.off()
```


# Differential abundance analysis with temperature on polyps
```{r, fig.width=10, fig.height=10, message=F, warning=F, eval=T}
library(EnhancedVolcano); library(tibble)

polyps <- subset_samples(ps_seasonal, Type=="polyp")
polypsM25 <- subset_samples(polyps, PopID=="Deep lake")
polypsM26 <- subset_samples(polyps, PopID=="River")
polypsM52 <- subset_samples(polyps, PopID=="Shallow lake")

load("/storage/microbiome/seasonal_2024/polyp_ancom_results.Rdata")
#temp.ancom <- ancombc2(polyps, fix_formula = "Temp", tax_level="Genus",p_adj_method = "BH")
#polyp.temp.ancom.M25 <- ancombc2(polypsM25, fix_formula = "Temp", tax_level="Genus",p_adj_method = "BH")
#polyp.temp.ancom.M26 <- ancombc2(polypsM26, fix_formula = "Temp", tax_level="Genus",p_adj_method = "BH")
#polyp.temp.ancom.M52 <- ancombc2(polypsM52, fix_formula = "Temp", tax_level="Genus",p_adj_method = "BH")
save(polyp.temp.ancom.M25, polyp.temp.ancom.M26, polyp.temp.ancom.M52,
     file="/storage/microbiome/seasonal_2024/polyp_ancom_results.Rdata")

daa.M25 <- polyp.temp.ancom.M25$res%>%column_to_rownames("taxon")
daa.M26 <- polyp.temp.ancom.M26$res%>%column_to_rownames("taxon")
daa.M52 <- polyp.temp.ancom.M52$res%>%column_to_rownames("taxon")

keyvals.colour.M52 <- ifelse(
    daa.M52$lfc_Temp < -0.05 & daa.M52$q_Temp<0.01, 'royalblue',
      ifelse(daa.M52$lfc_Temp > 0.05 & daa.M52$q_Temp<0.01, 'red2',
        'black'))
keyvals.colour.M52[is.na(keyvals.colour.M52)] <- 'black'
names(keyvals.colour.M52)[keyvals.colour.M52 == 'red2'] <- 'warm-specific'
names(keyvals.colour.M52)[keyvals.colour.M52 == 'black'] <- 'non-specific'
names(keyvals.colour.M52)[keyvals.colour.M52 == 'royalblue'] <- 'cold-specific'

keyvals.colour.M25 <- ifelse(
    daa.M25$lfc_Temp < -0.05 & daa.M25$q_Temp<0.01, 'royalblue',
      ifelse(daa.M25$lfc_Temp > 0.05 & daa.M25$q_Temp<0.01, 'red2',
        'black'))
keyvals.colour.M25[is.na(keyvals.colour.M25)] <- 'black'
names(keyvals.colour.M25)[keyvals.colour.M25 == 'red2'] <- 'warm-specific'
names(keyvals.colour.M25)[keyvals.colour.M25 == 'black'] <- 'non-specific'
names(keyvals.colour.M25)[keyvals.colour.M25 == 'royalblue'] <- 'cold-specific'

keyvals.colour.M26 <- ifelse(
    daa.M26$lfc_Temp < -0.05 & daa.M26$q_Temp<0.01, 'royalblue',
      ifelse(daa.M26$lfc_Temp > 0.05 & daa.M26$q_Temp<0.01, 'red2',
        'black'))
keyvals.colour.M26[is.na(keyvals.colour.M26)] <- 'black'
names(keyvals.colour.M26)[keyvals.colour.M26 == 'red2'] <- 'warm-specific'
names(keyvals.colour.M26)[keyvals.colour.M26 == 'black'] <- 'non-specific'
names(keyvals.colour.M26)[keyvals.colour.M26 == 'royalblue'] <- 'cold-specific'

polyp.v1 <- EnhancedVolcano(daa.M52, lab=rownames(daa.M52), x="lfc_Temp", y="q_Temp", 
                FCcutoff=0.05, pCutoff=0.01, drawConnectors = T,
                colCustom = keyvals.colour.M52,max.overlaps=15,
                xlab=bquote(~beta [Temperature]),labSize=4,
                title="Polyps - Shallow lake", subtitle=NULL,caption=NULL)+ xlim(-0.4,0.5)+ylim(0,8)
polyp.v2 <- EnhancedVolcano(daa.M25, lab=rownames(daa.M25), x="lfc_Temp", y="q_Temp", 
                FCcutoff=0.05, pCutoff=0.01, drawConnectors = T,
                colCustom = keyvals.colour.M25,max.overlaps=15,
                xlab=bquote(~beta [Temperature]),labSize=4,
                title="Polyps - Deep lake",subtitle=NULL,caption=NULL)+ xlim(-0.4,0.5)+ylim(0,8)
polyp.v3 <- EnhancedVolcano(daa.M26, lab=rownames(daa.M26), x="lfc_Temp", y="q_Temp", 
                FCcutoff=0.05, pCutoff=0.01, drawConnectors = T,
                colCustom = keyvals.colour.M26,max.overlaps=15,
                xlab=bquote(~beta [Temperature]),labSize=4,
                title="Polyps - River",subtitle=NULL,caption=NULL)+xlim(-0.4,0.5)+ylim(0,8)

ragg::agg_tiff(filename="/storage/microbiome/seasonal_2024/figs/polyp_daa.tiff",
               width=6, height=8,units="in",res=300, scaling=0.6)
ggarrange(polyp.v1+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          polyp.v2+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          polyp.v3+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"), ncol=1,common.legend = T)
dev.off()
```

# Differential abundance analysis with temperature on water
```{r, fig.width=10, fig.height=10, message=F, warning=F, eval=T}
library(EnhancedVolcano); library(tibble)

water <- subset_samples(ps_seasonal, Type=="water")
waterM25 <- subset_samples(water, PopID=="Deep lake")
waterM26 <- subset_samples(water, PopID=="River")
waterM52 <- subset_samples(water, PopID=="Shallow lake")

load("/storage/microbiome/seasonal_2024/water_ancom_results.Rdata")
#temp.ancom <- ancombc2(polyps, fix_formula = "Temp", tax_level="Genus",p_adj_method = "BH")
#water.temp.ancom.M25 <- ancombc2(waterM25, fix_formula = "Temp", tax_level="Genus",p_adj_method = "BH")
#water.temp.ancom.M26 <- ancombc2(waterM26, fix_formula = "Temp", tax_level="Genus",p_adj_method = "BH")
#water.temp.ancom.M52 <- ancombc2(waterM52, fix_formula = "Temp", tax_level="Genus",p_adj_method = "BH")
save(water.temp.ancom.M25, water.temp.ancom.M26, water.temp.ancom.M52,
     file="/storage/microbiome/seasonal_2024/water_ancom_results.Rdata")

daa.M25 <- water.temp.ancom.M25$res%>%column_to_rownames("taxon")
daa.M26 <- water.temp.ancom.M26$res%>%column_to_rownames("taxon")
daa.M52 <- water.temp.ancom.M52$res%>%column_to_rownames("taxon")

keyvals.colour.M52 <- ifelse(
    daa.M52$lfc_Temp < -0.05 & daa.M52$q_Temp<0.01, 'royalblue',
      ifelse(daa.M52$lfc_Temp > 0.05 & daa.M52$q_Temp<0.01, 'red2',
        'black'))
keyvals.colour.M52[is.na(keyvals.colour.M52)] <- 'black'
names(keyvals.colour.M52)[keyvals.colour.M52 == 'red2'] <- 'warm-specific'
names(keyvals.colour.M52)[keyvals.colour.M52 == 'black'] <- 'non-specific'
names(keyvals.colour.M52)[keyvals.colour.M52 == 'royalblue'] <- 'cold-specific'

keyvals.colour.M25 <- ifelse(
    daa.M25$lfc_Temp < -0.05 & daa.M25$q_Temp<0.01, 'royalblue',
      ifelse(daa.M25$lfc_Temp > 0.05 & daa.M25$q_Temp<0.01, 'red2',
        'black'))
keyvals.colour.M25[is.na(keyvals.colour.M25)] <- 'black'
names(keyvals.colour.M25)[keyvals.colour.M25 == 'red2'] <- 'warm-specific'
names(keyvals.colour.M25)[keyvals.colour.M25 == 'black'] <- 'non-specific'
names(keyvals.colour.M25)[keyvals.colour.M25 == 'royalblue'] <- 'cold-specific'

keyvals.colour.M26 <- ifelse(
    daa.M26$lfc_Temp < -0.05 & daa.M26$q_Temp<0.01, 'royalblue',
      ifelse(daa.M26$lfc_Temp > 0.05 & daa.M26$q_Temp<0.01, 'red2',
        'black'))
keyvals.colour.M26[is.na(keyvals.colour.M26)] <- 'black'
names(keyvals.colour.M26)[keyvals.colour.M26 == 'red2'] <- 'warm-specific'
names(keyvals.colour.M26)[keyvals.colour.M26 == 'black'] <- 'non-specific'
names(keyvals.colour.M26)[keyvals.colour.M26 == 'royalblue'] <- 'cold-specific'

water.v1 <- EnhancedVolcano(daa.M52, lab=rownames(daa.M52), x="lfc_Temp", y="q_Temp", 
                FCcutoff=0.05, pCutoff=0.01, drawConnectors = T,
                colCustom = keyvals.colour.M52,max.overlaps=15,
                xlab=bquote(~beta [Temperature]),labSize=4,
                title="Water - Shallow lake",subtitle=NULL,caption=NULL)+ xlim(-0.4,0.5)+ylim(0,8)
water.v2 <- EnhancedVolcano(daa.M25, lab=rownames(daa.M25), x="lfc_Temp", y="q_Temp", 
                FCcutoff=0.05, pCutoff=0.01, drawConnectors = T,
                colCustom = keyvals.colour.M25,max.overlaps=15,
                xlab=bquote(~beta [Temperature]),labSize=4,
                title="Water - Deep lake",subtitle=NULL,caption=NULL)+ xlim(-0.4,0.5)+ylim(0,8)
water.v3 <- EnhancedVolcano(daa.M26, lab=rownames(daa.M26), x="lfc_Temp", y="q_Temp", 
                FCcutoff=0.05, pCutoff=0.01, drawConnectors = T,
                colCustom = keyvals.colour.M26,max.overlaps=15,
                xlab=bquote(~beta [Temperature]),labSize=4,
                title="Water - River",subtitle=NULL,caption=NULL)+xlim(-0.4,0.5)+ylim(0,8)

ragg::agg_tiff(filename="/storage/microbiome/seasonal_2024/figs/water_daa.tiff",
               width=6, height=8,units="in",res=300, scaling=0.6)
ggarrange(water.v1+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          water.v2+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          water.v3+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"), ncol=1,common.legend = T)
dev.off()
```

```{r}
ragg::agg_tiff(filename="/storage/microbiome/seasonal_2024/figs/daa.tiff",
               width=10, height=8,units="in",res=300, scaling=0.6)
ggarrange(polyp.v1+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          water.v1+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          polyp.v2+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          water.v2+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          polyp.v3+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"),
          water.v3+theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
                   legend.position = "top"), ncol=2,nrow=3,common.legend = T)
dev.off()
```

#Stochasticity of microbial communities with NST
```{r, fig.width=8, fig.height=12, eval=F}
library(NST)
comm.M25 <- t(otu_table(polypsM25))
group.M25 <- sample_data(polypsM25)%>%data.frame()%>%select("Sampling")

nst.M25 <- tNST(comm.M25, group.M25, meta.group=NULL, meta.com=NULL,
          dist.method="bray", abundance.weighted=TRUE, rand=1000,
          output.rand=FALSE, nworker=8, LB=FALSE, null.model="PF",
          between.group=TRUE, SES=TRUE, RC=TRUE)
nst.M25.between <- nst.M25$index.between%>%filter(group%in%c("2021 June.vs.2021 September",
                                          "2021 September.vs.2021 December",
                                          "2021 December.vs.2022 March",
                                          "2022 March.vs.2022 June",
                                          "2022 June.vs.2022 November",
                                          "2022 November.vs.2023 February",
                                          "2023 February.vs.2023 April",
                                          "2023 April.vs.2023 June"))
nst.M25.between$group <- factor(nst.M25.between$group, 
                                levels=unique(nst.M25.between$group))
nst.plot.M25 <- ggplot(nst.M25.between, aes(x=group, y=NST.i.bray*100, group=1))+geom_point()+geom_line()+
  theme_bw()+geom_hline(yintercept=50, linetype="dashed", color="blue")+
  ylim(0,100)+ylab("NST (%)")+labs(title="Population M25")+
  annotate("text", x=1.5, y =90, label = "Stochastic processes", vjust = 0, color = "blue")+
  annotate("text", x=1.5, y = 5, label = "Deterministic processes", vjust = 0, color = "blue") +
  theme(axis.text.x = element_text(angle=45,hjust=1))

comm.M26 <- t(otu_table(polypsM26))
group.M26 <- sample_data(polypsM26)%>%data.frame()%>%select("Sampling")

nst.M26 <- tNST(comm.M26, group.M26, meta.group=NULL, meta.com=NULL,
          dist.method="bray", abundance.weighted=TRUE, rand=1000,
          output.rand=FALSE, nworker=8, LB=FALSE, null.model="PF",
          between.group=TRUE, SES=TRUE, RC=TRUE)
nst.M26.between <- nst.M26$index.between%>%filter(group%in%c("2021 June.vs.2021 September",
                                          "2021 September.vs.2021 December",
                                          "2021 December.vs.2022 March",
                                          "2022 March.vs.2022 June",
                                          "2022 June.vs.2022 November",
                                          "2022 November.vs.2023 February",
                                          "2023 February.vs.2023 April",
                                          "2023 April.vs.2023 June"))
nst.M26.between$group <- factor(nst.M26.between$group, 
                                levels=unique(nst.M26.between$group))
nst.plot.M26 <- ggplot(nst.M26.between, aes(x=group, y=NST.i.bray*100, group=1))+geom_point()+geom_line()+
  theme_bw()+geom_hline(yintercept=50, linetype="dashed", color="blue")+
  ylim(0,100)+ylab("NST (%)")+labs(title="Population M26")+
  annotate("text", x=1.5, y = 90, label = "Stochastic processes", vjust = 0, color = "blue")+
  annotate("text", x=1.5, y = 5, label = "Deterministic processes", vjust = 0, color = "blue") +
  theme(axis.text.x = element_text(angle=45,hjust=1))

comm.M52 <- t(otu_table(polypsM52))
group.M52 <- sample_data(polypsM52)%>%data.frame()%>%select("Sampling")

nst.M52 <- tNST(comm.M52, group.M52, meta.group=NULL, meta.com=NULL,
          dist.method="bray", abundance.weighted=TRUE, rand=1000,
          output.rand=TRUE, nworker=8, LB=FALSE, null.model="PF",
          between.group=TRUE, SES=TRUE, RC=TRUE)
nst.M52.between <- nst.M52$index.between%>%filter(group%in%c("2021 June.vs.2021 September",
                                          "2021 September.vs.2021 December",
                                          "2021 December.vs.2022 March",
                                          "2022 March.vs.2022 June",
                                          "2022 June.vs.2022 November",
                                          "2022 November.vs.2023 February",
                                          "2023 February.vs.2023 April",
                                          "2023 April.vs.2023 June"))
nst.M52.between$group <- factor(nst.M52.between$group, 
                                levels=unique(nst.M52.between$group))
nst.plot.M52 <- ggplot(nst.M52.between, aes(x=group, y=NST.i.bray*100, group=1))+geom_point()+geom_line()+
  theme_bw()+geom_hline(yintercept=50, linetype="dashed", color="blue")+
  ylim(0,100)+ylab("NST (%)")+labs(title="Population M52")+
  annotate("text", x=1.5, y = 90, label = "Stochastic processes", vjust = 0, color = "blue")+
  annotate("text", x=1.5, y = 5, label = "Deterministic processes", vjust = 0, color = "blue") +
  theme(axis.text.x = element_text(angle=45,hjust=1))

ggpubr::ggarrange(nst.plot.M25, nst.plot.M26, nst.plot.M52, ncol=1)
```


```{r, fig.width=10, message=T, warning=T, eval=F}
#samplings <- unique(sample_data(ps_seasonal)$Sampling)
#props_spring <- list()
#for(i in 1:length(samplings)){
#  print(i)
#  ps_sel <- subset_samples(ps_seasonal,Sampling==samplings[i]&Type=="polyp")

  ps_sel <- subset_samples(ps_seasonal, Type=="polyp")
  taxa_names(ps_sel) <- tax_table(ps_sel)[,6]

  sel_tax <- top_taxa(subset_samples(ps_sel, Type=="polyp"),n=50)$top_taxa[,2]

  ps_sel <- subset_taxa(ps_sel, Genus%in%sel_tax)
  print(ps_sel)

  net_spring <- netConstruct(data = ps_sel,
                           measure = "spring",
                           measurePar = list(nlambda=50, 
                                             rep.num=50),
                           normMethod = "none", 
                           zeroMethod = "none",
                           sparsMethod = "none", 
                           dissFunc = "signed",
                           verbose = 2,
                           seed = 123456)

  props_spring <- netAnalyze(net_spring, 
                           centrLCC = TRUE,
                           clustMethod = "cluster_fast_greedy",
                           hubPar = "eigenvector",
                           weightDeg = FALSE, normDeg = FALSE)
#}

p <- plot(props_spring, 
          nodeColor = "cluster", 
          nodeSize = "eigenvector",
          nodeSizeSpread = 2,
          cexNodes = 4,
          showTitle = FALSE,
          labelScale = TRUE,
          cexLabels = 5)

save(net_spring, props_spring, file="/storage/microbiome/seasonal_2024/net100.Rdata")

```


```{r, fig.width=10, fig.height=10, message=F, warning=F, eval=F}
library(igraph); library(tidygraph); library(ggraph); library(ggforce); 
library(scatterpie); library(grafify); library(ggnewscale)

load("/storage/microbiome/seasonal_2024/net100.Rdata")

plot_graph <- as_tbl_graph(graph_from_adjacency_matrix(net_spring$adjaMat1,weighted=T,mode="undirected"))

node.list <- plot_graph%>%activate(nodes)%>%data.frame()%>%as.vector()
node.data <- data.frame(node=node.list$name,
                        Cluster=as.factor(props_spring$clustering$clust1[node.list$name]),
                        polyp.abund=taxa_sums(subset_samples(ps_sel, Type=="polyp"))[node.list$name],
                        spring.abund=taxa_sums(subset_samples(ps_sel, Season=="Spring"))[node.list$name],
                        summer.abund=taxa_sums(subset_samples(ps_sel, Season=="Summer"))[node.list$name],
                        autumn.abund=taxa_sums(subset_samples(ps_sel, Season=="Autumn"))[node.list$name],
                        winter.abund=taxa_sums(subset_samples(ps_sel, Season=="Winter"))[node.list$name])
                        
plot_graph <- plot_graph%>%activate(nodes)%>%left_join(node.data, by=c("name"="node"))

V(plot_graph)$p <- taxa_sums(subset_samples(ps_sel, Type=="polyp"))[node.list$name]
V(plot_graph)$grp <- props_spring$clustering$clust1[node.list$name]

set.seed(123)
xy <- layout_with_kk(plot_graph)#, weights=0.1*E(plot_graph2)$weight)
V(plot_graph)$x <- xy[,1]
V(plot_graph)$y <- xy[,2]
V(plot_graph)$Cluster <- as.factor(V(plot_graph)$grp)
plot_graph2 <- as_tbl_graph(plot_graph)

net <- ggraph(plot_graph2, "manual", x = V(plot_graph)$x, y = V(plot_graph)$y)+
  geom_edge_fan(color="grey90")+theme_graph()+
#  geom_mark_hull(aes(x,y,group=Cluster,fill=Cluster),
#                       concavity=8,alpha=0.25,linetype=2)+
  scale_fill_grafify(name="Cluster:",palette="fishy")+
  scale_color_grafify(name="Cluster:",palette="fishy")+
#  new_scale_fill()+
  geom_node_point(aes(color=Cluster, size=log10(polyp.abund)))+
  scale_size_continuous(range=c(1,10))+    
#  scale_fill_manual(name="Type:",values=c("w"="lightblue3","p"="orange"),
#                    labels=c("w"="water","p"="polyp"))+
  geom_node_text(aes(label=ifelse(polyp.abund > quantile(polyp.abund,0.5), 
                                  as.character(name), NA_character_),
                    # size=2*log10(polyp.abund)
                     ),
                 repel=T,force=25,force_pull=0)+guides(size="none")
net
```
